\bfseries Emulating cosmological n-body simulations with equivariant convolutional neural networks

### 2022-05-19 paper/main.tex

\begin{definition}[vectors, $k$-tensors, and the world]
Let $O(d)$ act on vectors in $\mathbb R^d$ via the canonical action.
It also acts on lists of vectors $(\mathbb R^d)^n$ by acting on each of the $n$ vectors by the same action.
It also acts on rank-1 $k$-tensors in $(\mathbb R^d)^{\otimes k}$ by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = g\cdot v_1\otimes \ldots \otimes g\cdot v_k$ and extending linearly to higher-rank tensors made from sums of rank-1 tensors.
Let $W$ (aka ``the world'') be the comprehensive set of all vectors and tensors that exist.
Note that $O(d)$ can be seen as acting on the world.
\end{definition}

@@

\begin{definition}[linear geometric convolution]
Given an image $A$ of geometric objects with $N^d$ pixels, we define its linear convolution with a geometric filter $C$ as follows.
Each pixel $p$ (numbered $1\leq p\leq N^d$) of the image contains a geometric object $a_p$.
All of these $N^d$ geometric objects $a_p$ are of the same geometric type ($k$-tensor or $k$-pseudotensor).
The filter $C$ has side length $M\leq N$, where $M$ is an odd integer, such that it has $M^d$ total pixels.
Each pixel $m$ (numbered $1\leq m\leq M^d$) of the filter contains a geometric object $w_m$.
All of these $M^d$ geometric weights $w_m$ are of the same geometric type as well ($k'$-tensor or $k'$-pseudotensor).
For image pixel $p$ (with $1\leq p\leq N^d$) we take a $M^d$ patch of the image, centered at $p$ with contents $a(p)$ (where we assume that the image is in the $d$-torus, so every pixel centers a $M^d$ patch) and number its pixels $1\leq m\leq M^d$.
Then the convolution $A\star C$ of the image $A$, evaluated at the central pixel $p$ of the patch $P_p$ (containing pixel values $a(p)_m$ with $1\leq m\leq M^d$) and the filter $C$ is the sum of outer products
\begin{equation}
    (A\star C)_p := \sum_{m=1}^{M^d} a(p)_m \otimes w_m.
\end{equation}
\end{definition}

@@

Contractions of these objects are possible, including contractions that involve either the (even parity) Kroenecker delta $\delta_{ij}$ or the (odd parity) Levi-Civita symbol $\epsilon_{ijk}$ (in $d=3$ dimensions; $\epsilon_{ij}$ in $d=2$ dimensions; $\epsilon_{ijk\ell}$ in $d=4$ dimensions).
The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices such that if the $d$ indices are not repeated and in cyclic order the value is $+1$ and if the $d$ indices are not repeated and in anti-cyclic order the value is $-1$, and it has the value $0$ in all other cases.
For an example of how the Levi-Civita symbol is used, in $d=3$ dimensions, the determinant of a $d\times d$ matrix $A$ is given by $\det(A) = [A]_{1,i}\,[A]_{2,j}\,[A]_{3,j}\,\epsilon_{ijk}$ in summation notation.
Contraction of a $k$-tensor (with $k\geq 2$) with one application of the Kroenecker delta returns a $(k-2)$-tensor.
Contraction of a $k$-pseudotensor (with $k\geq 2$) with one application of the Kroenecker delta returns a $(k-2)$-pseudotensor where $d$ is the spatial dimension.
Contraction of a $k$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-pseudotensor, because the contractions are structured with the last of the $d$ indices of the Levi-Civita symbol unpaired. SOLE DO WE NEED TO SAY MORE HERE?
Contraction of a $k$-pseudotensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-tensor.

For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $[a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
For another example, the same 2-tensor can be contracted with the Levi-Civita symbol to give the standard cross product
$[a]_i\,[b]_j\,\epsilon_{ijk}$, which is a 1-pseudotensor or pseudovector.

@@

\section{Equivariant linear geometric convolution operators}

@@

\begin{definition}[Equivariant convolution operator]
Foo and bar.
\end{definition}

@@

HOGG: What is a linear convolution operator?

HOGG: What is a linear geometric convolution operator?

@@

\section{Equivariant nonlinear geometric convolutions}

\section{Construction of all equivariant convolution operators}

>>>>

\begin{definition}[vectors, $k$-tensors, and the world]
Let $O(d)$ act on vectors in $\mathbb R^d$ via the canonical action.
It also acts on lists of vectors $(\mathbb R^d)^n$ by acting on each of the $n$ vectors by the same action.
It also acts on rank-1 $k$-tensors in $(\mathbb R^d)^{\otimes k}$ by $g\cdot (v_{1}\otimes\ldots \otimes v_k) = (g\cdot v_1)\otimes \ldots \otimes (g\cdot v_k)$ and extending linearly to higher-rank tensors made from sums of rank-1 tensors.
Let $W$ (aka ``the world'') be the comprehensive set of all vectors and tensors that exist.
Note that $O(d)$ can be seen as acting on the world.
\end{definition}

@@

HOGG: Define a geometric image here.

HOGG: Define a geometric filter here. Then shorten the defition below.

\begin{definition}[geometric convolution, geometric convolution filter]
Let $A$ be a $d$-dimensional image of $N^d$ pixels, in each of which there is a $k$-tensor or $k$-pseudotensor as defined above.
The image $A$ lives in a cubical $d$-torus such that the image has no boundary.
Let $C$ be a $d$-dimensional image with $M^d$ pixels (with $M$ odd and $M<N$), in each of which there is a $k'$-tensor or $k'$-pseudotensor.
We call $C$ a geometric convolution filter.
The geometric convolution $A\star C$ is a $d$-dimensional image of $N^d$ pixels, in each of which there is a $k+k'$-tensor or $k+k'$-pseudotensor, defined as follows:
For image pixel $p$ in $A\star C$ (with $1\leq p\leq N^d$) we take the cubical $M^d$ patch of the image centered on pixel $p$ with contents $a(p)$ (the image is on the $d$-torus, so every pixel $p$ centers a $M^d$ patch) and number its pixels $1\leq m\leq M^d$.
Then the convolution $A\star C$ of the image $A$, evaluated at the central pixel $p$ of the cubical patch (containing pixel values $a(p)_m$ with $1\leq m\leq M^d$) and the filter $C$ is the sum of outer products
\begin{equation}
    (A\star C)_p := \sum_{m=1}^{M^d} a(p)_m \otimes w_m~.
\end{equation}
That is, geometric convolution is the same as ordinary convolution on the $d$-torus except generalized to outer products of tensor and pseudotensor inputs and outputs.
\end{definition}

@@

THIS PARAGRAPH SUX: Contractions of these objects are possible, including contractions that involve either the (even parity) Kroenecker delta $\delta_{ij}$ or the (odd parity) Levi-Civita symbol $\epsilon_{ijk}$ (in $d=3$ dimensions; $\epsilon_{ij}$ in $d=2$ dimensions; $\epsilon_{ijk\ell}$ in $d=4$ dimensions).
The Levi-Civita symbol in $d\geq 2$ spatial dimensions is the object with $d$ indices such that if the $d$ indices are not repeated and in cyclic order the value is $+1$ and if the $d$ indices are not repeated and in anti-cyclic order the value is $-1$, and it has the value $0$ in all other cases.
For an example of how the Levi-Civita symbol is used, in $d=3$ dimensions, the determinant of a $d\times d$ matrix $A$ is given by $\det(A) = [A]_{1,i}\,[A]_{2,j}\,[A]_{3,j}\,\epsilon_{ijk}$ in summation notation.
Contraction of a $k$-tensor (with $k\geq 2$) with one application of the Kroenecker delta returns a $(k-2)$-tensor.
Contraction of a $k$-pseudotensor (with $k\geq 2$) with one application of the Kroenecker delta returns a $(k-2)$-pseudotensor where $d$ is the spatial dimension.
Contraction of a $k$-tensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-pseudotensor, because the contractions are structured with the last of the $d$ indices of the Levi-Civita symbol unpaired. SOLE DO WE NEED TO SAY MORE HERE?
Contraction of a $k$-pseudotensor (with $k\geq (d-1)$) with one application of the Levi-Civita symbol returns a $(k-d+2)$-tensor.

For one example of these contractions, the 2-tensor formed by the outer product of 1-tensors $a$ and $b$ can be contracted with the Kronecker delta to give the standard dot product $a^\top b = [a]_i\,[b]_j\,\delta_{ij}$, which is a 0-tensor or scalar.
For another example, the same 2-tensor can be contracted with the Levi-Civita symbol to give the standard cross product
$[a\times b]_k = [a]_i\,[b]_j\,\epsilon_{ijk}$, which is a 1-pseudotensor or pseudovector.

@@

\section{Equivariant geometric convolution operators}

@@

\begin{definition}[Group $B_d$ of symmetries of a $d$- hypercube]
We denote by $B_d$ the group of rotation and reflection symmetries of the $d$-dimensional hypercube.
\end{definition}

Notes about the group $B_d$: In $d=2$ there are $8$ group elements, in $d=3$ there are $48$ group elements (and in $d=4$ there are 384). Because the groups $B_d$ are subgroups of $O(d)$, all determinants of the matrix representations of the group elements are either $+1$ or $-1$, and the matrix representation $M(g^{-1})$ of the inverse $g^{-1}$ of group element $g$ is the transpose of the matrix representation $M(g)$ of group element $g$.

\begin{definition}[Action of $B_d$ on $k$-tensors and $k$-pseudotensors]
\begin{align}
\mbox{$k$-tensor:} ~~ g\cdot(v_1 \otimes \ldots \otimes v_k) &= (g\cdot v_1)\otimes\ldots\otimes(g\cdot v_k) \\
\mbox{$k$-pseudotensor:} ~~ g\cdot(v_1 \otimes \ldots \otimes v_k) &= \det(g)\,(g\cdot v_1)\otimes\ldots\otimes(g\cdot v_k) ~,
\end{align}
And extend linearly
\end{definition}

This may seem familiar: The $k$-tensor and $k$-pseudotensor were defined by how they transform under the orthogonal group $O(d)$, and our group $B_d$ is a subgroup of the orthogonal group.

\begin{definition}[Action of $B_d$ on geometric images]
Given $A$ a cubic $N^d$ image with a $k$-tensor in each pixel, and $g\in B_d$ we define $g\cdot A$ a $N^d$-dimensional image with $k$-tensors in each pixel such that
\begin{equation}
    (g\cdot A)_p = g\cdot A_{g^{-1}\cdot p}
\end{equation}
\end{definition}

\begin{definition}[The total group $G_d$ in play]
Outer product of the above group and discrete translations on the $d$-cubic lattice.
\end{definition}

\begin{definition}[Equivariant convolution operator]
A geometric convolution filter $C$ is equivariant if it is an equivariant function with respect to the action above. 
\end{definition}

@@

@@

\section{Nonlinear equivariant local geometric operators}

\section{Construction of all equivariant operators}
